Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-04-11 GWLDCI.tex
fixed notation

The first conditions states that there are at least four more vertices than propagators in an admissible Wilson Loop Diagram. The second imposes an upper bound on how densely the propagators can be fitted in the diagram. The third ensures that ensures that no propagators cross in the interior of the diagram. In other words, a Wilson loop diagram, $(\cP, [n])$ is admissible if and only if $n > \cP +4$, and has neither crossing propagators nor any pairs of propagators that start and end on the same pair of non-adjacent edges.

@@

\begin{dfn}\label{equivdfn} 
There is an equivalence relationship on the set of admissible Wilson loops diagrams given by the transititve closure the following binary relation: $W = (\cP, [n]) \sim W'= (\cP', n)$ if
\begin{enumerate}
\item There exist two different exact subdiagrams, $(P, V(P))$ and $(P', V(P'))$ of $W$ and $W'$ respectively such that $V(P) =  V(P')$.
\item The complementary subdiagrams are identical: $(\cP \setminus P, V(P)^c) = (\cP' \setminus P', V(P')^c)$.
\end{enumerate}
\end{dfn}

@@

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note that since vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$, $M(W)$ can also be thought of as the matroid realized by $C(W)$. \note{Can this note have a ref to your earlier paper?  If not then perhaps expand slightly as although I agree that it is true, I don't think it is quite so trivial.  Also, I think ``is also'' woulc be fine in place of ``can also be thought of as''}

@@

\begin{thm}\label{same matroid iff equiv}
Let $W= (\cP, [n])$ and $W'= (\cP', [n])$ be two Wilson Loop diagrams. They define the same matroid if and only if $W \sim W'$.
\end{thm}

@@

The map $\tau$ defined in Definition \ref{WLDtriangulationdfn} assigns a dissection of an $n$-gon to a Wilson loop diagram $W = (\cP, n)$. Recall that if $|\cP| = k$, then $\tau(W)$ is a dissection with $k$ chords. Therefore, the map $\tau$ associates to each weakly admissible Wilson loop diagram a $n-3 - k$ dimensional face of the associahedron $A_n$. \sanote{picture of $\tau(W)$ for $k =1$, $n = 5$?}

>>>>

The first conditions states that there are at least four more vertices than propagators in an admissible Wilson loop diagram. The second imposes an upper bound on how densely the propagators can be fitted in the diagram. The third ensures that ensures that no propagators cross in the interior of the diagram. \note{fixed? earlier, we had $i + 1 < j$ in the definition. When that was fixed, I didn't propagate forward. Or am I missing something else?}In other words, a Wilson loop diagram, $(\cP, [n])$ is admissible if and only if $n \geq \cP +4$, and has no crossing propagators, no propagators of the form $p = (i, i+1)$, and no pairs of propagators that start and end on the same pair of non-adjacent edges.

@@

\begin{dfn}\label{equivdfn} 
There is an equivalence relationship on the set of admissible Wilson loops diagrams given by the transititve closure the following binary relation: $W = (\cP, [n]) \sim W'= (\cP', [n])$ if
\begin{enumerate}
\item There exist two different exact subdiagrams, $(P, V(P))$ and $(P', V(P'))$ of $W$ and $W'$ respectively such that $V(P) =  V(P')$.
\item The complementary subdiagrams are identical: $(\cP \setminus P, V(P)^c) = (\cP' \setminus P', V(P')^c)$.
\end{enumerate}
\end{dfn}

@@

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note from equation \eqref{C(W) dfn}that the vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$. Therefore, for any subset of vertices $V \subset [n]$,  then one can write $C(W)|_V$ is the restriction of $C(W)$ to the columns corresponding to $V$. Since the non-zero entries of $C(W)$ are real variables, the matrix $C(W)|_V$ does not have full rank if and only if there are fewer rows with non-zero entries than $|V|$. This is equivalent to saying that $V$ is a set of dependent columns of $C(W)$ if and only if $|\Prop(V)| < |V|$. Comparing this to Theorem \ref{thm:WLDmatroid}, we see that $M(W)$ is also the matroid realized by $C(W)$. \sanote{Is this suffient for an explanation?}

@@

\begin{thm}\label{same matroid iff equiv}
Let $W= (\cP, [n])$ and $W'= (\cP', n)$ be two Wilson Loop diagrams. They define the same matroid if and only if $W \sim W'$.
\end{thm}

@@

The map $\tau$ defined in Definition \ref{WLDtriangulationdfn} assigns a dissection of an $n$-gon to a Wilson loop diagram $W = (\cP, [n])$. Recall that if $|\cP| = k$, then $\tau(W)$ is a dissection with $k$ chords. Therefore, the map $\tau$ associates to each weakly admissible Wilson loop diagram a $n-3 - k$ dimensional face of the associahedron $A_n$. \sanote{picture of $\tau(W)$ for $k =1$, $n = 5$?}
